---
title: "scRNA_seq"
format: html
editor: visual
toc: true
---

# Set up 
```{r}
library(here)
library(glue)
library(Seurat)
library(ggplot2)
library(dplyr)
library(HGNChelper)
library(openxlsx)
source(here("utils/sctype_wrapper.R"))


dir <- "scRNA_seq/"

# Create directories
directories <- c(here(dir, "rds"))
sapply(directories, function(x){dir.create(x, showWarnings = FALSE, recursive = TRUE)})

```

## Process: QC to UMAP

The aim is to process and integrate WT, MT and ZK samples, following Seurat vignette.

```{r}
# Name and path of datasets dir
list_seurat <- list(
  # WT
  "WT" = list(
    condition = "WT",
    data.dir = here(dir, "data/S22096/MNGC484_filtered_feature_bc_matrix/")),
  # Cyp24a1 null 10 weeks
  "MT" = list(
    condition = "MT",
    data.dir = here(dir, "data/S22096/MNGC485_filtered_feature_bc_matrix/")),
  # ZK-treated Cyp24a1 null 10 weeks
  "ZK" = list(
    condition = "ZK",
    data.dir = here(dir, "data/S22096/MNGC486_filtered_feature_bc_matrix/"))
)

# Creation of Seurat objects
list_seurat <- lapply(list_seurat, function(x){
  # Load the data
  data <- Read10X(data.dir = x$data.dir)
  seurat <- CreateSeuratObject(data,
                               project = x$condition,
                               assay = "RNA",
                               min.cells = 10,
                               min.features = 100,
                               names.field = 1,
                               names.delim = "_",
                               meta.data = NULL
  )
  # Add percent.mt feature
  feature <- list(col.name = "percent.mt",
                  pattern = "^MT-|^mt-|^Mt-")

  seurat <- PercentageFeatureSet(object = seurat,
                                 col.name = feature$col.name,
                                 pattern = feature$pattern)

  feature <- list(col.name = "percent.ribo",
                  pattern = "^Rp[sl][[:digit:]]|^Rp[[:digit:]]|^Rpsa")

  seurat <- PercentageFeatureSet(object = seurat,
                                 col.name = feature$col.name,
                                 pattern = feature$pattern)

  # Save VlnPlot of unfiltered data
  VlnPlot(seurat, features = c("nFeature_RNA", "nCount_RNA", 
                               "percent.mt", "percent.ribo"),
          pt.size = 1, ncol = 3)

  # Filter data
  seurat <- subset(seurat, subset = nFeature_RNA > 100 &
                     nFeature_RNA < 7000 &
                     percent.mt < 50)

  # Save VlnPlot of filtered data
  VlnPlot(seurat, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.ribo"),
          pt.size = 1, ncol = 3)

  return(seurat)
})

# Integration of WT, MT and ZK objects

# Creation of the combined Seurat object
seurat <- merge(x = list_seurat$WT,
                y = list_seurat[c("MT", "ZK")])
# Normalize each object
seurat <- NormalizeData(seurat)
seurat <- FindVariableFeatures(seurat,
                               selection.method = "vst",
                               nfeatures = 2000)

seurat <- ScaleData(seurat, verbose = F)
seurat <- RunPCA(seurat, verbose = F)
# Integrate the datasets
seurat <- IntegrateLayers(seurat, method = CCAIntegration,
                          orig = "pca", new.reduction = "integrated.cca",
                          verbose = T)
# Cluster the integrated data
seurat <- FindNeighbors(seurat, reduction = "integrated.cca", dims = 1:30)
seurat <- FindClusters(seurat, resolution =  0.8)

seurat <- RunUMAP(seurat, reduction = "integrated.cca",
                  dims = 1:30, return.model = T, verbose = F)

# Create common counts and data layers
# seurat <- JoinLayers(seurat)

DimPlot(seurat, group.by = "orig.ident", reduction = "umap")
DimPlot(seurat, split.by = "orig.ident", 
        group.by = "seurat_clusters", reduction = "umap")

DefaultAssay(seurat) <- "RNA"
saveRDS(seurat, here(dir, "rds", "WT_MT_ZK_seurat.rds"))
```

## Annotate

The aim is to annotate clusters, with ScType tool (https://www.nature.com/articles/s41467-022-28803-w), based on Novella-Rausell et al. kidney markers (https://doi.org/10.1016/j.isci.2023.106877)

```{r}
seurat <- readRDS(here(dir, "rds", "WT_MT_ZK_seurat.rds"))
tissue = "Kidney"

# Create marker table in sctype format 
create_markers_table <- TRUE # Do it only once
if (create_markers_table){
  markers_all <- read.xlsx(here(dir,"data/1-s2.0-S2589004223009549-mmc5.xlsx"))
  markers <- markers_all %>%
    group_by(Cell.Population) %>%
    slice_head(n = 10) %>%
    aggregate(Gene.symbol ~ Cell.Population,
              FUN = function(x) paste(x, collapse = ",")) %>%
    mutate(tissueType = "Kidney", .before = 1) %>%
    rename("cellName" = "Cell.Population",
           "geneSymbolmore1" = "Gene.symbol") %>%
    mutate(geneSymbolmore2 = NA,
           shortName = cellName,
           .after = "geneSymbolmore1")
  write.xlsx(markers, file = "data/kidney_db.xlsx",
             sheetName="Kidney", append=TRUE)

  markers_filtered <- markers %>%
    filter(!cellName %in% c("Asc-Vasa-Recta", "CD-Trans", "CNT", "CTAL",
                            "DC", "DCT-CNT", "Desc-Vasa-Recta",
                            "DTL", "DTL-ATL", "Fib",
                            "Glom-Endo", "LOH", "MD", "MTAL", "PEC",
                            "Vas-Afferens", "Vas-Efferens", "Unknown"))

  write.xlsx(markers_filtered, file = "data/kidney_db_filtered.xlsx",
             sheetName="Kidney", append=TRUE)
}

# Compute sctype annotation
dbname <- "kidney_db_filtered"
kidney_db <- here(dir, glue("data/{dbname}.xlsx"))

seurat <- run_sctype(seurat, assay = "RNA", scaled = TRUE,
                     known_tissue_type = tissue,
                     custom_marker_file = kidney_db,
                     name = dbname)
seurat@meta.data[[dbname]] <- seurat@meta.data[[dbname]] %>% as.factor()

DimPlot(seurat, group.by = dbname,
        split.by = "orig.ident", label = TRUE) +
  theme(legend.position = 'none')

# Number of cells per annotation
summary(seurat@meta.data[[dbname]] %>% as.factor())

# Remove Unknown and re-annotate
expr <- FetchData(seurat, vars = dbname)
seurat <- seurat[, which(expr != "Unknown")]

# Re-run seurat analysis
seurat <- RunPCA(seurat, npcs = 50, verbose = T)
seurat <- IntegrateLayers(object = seurat, method = CCAIntegration,
                        orig.reduction = "pca", new.reduction = "integrated.cca",
                        verbose = FALSE)

# Re-join layers after integration
seurat[["RNA"]] <- JoinLayers(seurat[["RNA"]])
# to Re-split run:
# seurat_kidneymarkers[["RNA"]] <- split(seurat_kidneymarkers[["RNA"]], f = seurat_kidneymarkers$orig.ident)

seurat <- FindNeighbors(seurat,reduction = "integrated.cca", dims = 1:30)
seurat <- FindClusters(seurat, resolution =  0.8)
seurat <- RunUMAP(seurat, dims = 1:30, reduction = "integrated.cca")

# Re-compute sctype annotation
seurat <- run_sctype(seurat, assay = "RNA",
                             scaled = TRUE,
                              known_tissue_type = tissue,
                              custom_marker_file = kidney_db,
                              name = glue("{dbname}_2"))
seurat@meta.data[[glue("{dbname}_2")]] <- seurat@meta.data[[glue("{dbname}_2")]] %>% as.factor()

# Number of cells per annotation
summary(seurat@meta.data[[glue("{dbname}_2")]] %>% as.factor())

# Remove Unknown 
expr <- FetchData(seurat, vars = glue("{dbname}_2"))
seurat <- seurat[, which(expr != "Unknown")]

DimPlot(seurat, group.by = glue("{dbname}_2"),
        split.by = "orig.ident", label = TRUE) +
  theme(legend.position = 'none')

# Save results
saveRDS(seurat,
        here(dir, "rds", "WT_MT_ZK_seurat.rds"))
```

## Cell type markers

The aim is to compute cell type markers.

```{r}
sapply(c(here(dir, "table", "deg"),
         here(dir, "table", "pathway")),
       FUN = function(x){dir.create(x, recursive = T)})

dbname <- "kidney_db_filtered"
seurat <- readRDS(here(dir, "rds", glue("WT_MT_ZK_seurat.rds")))

seurat$CellType <- seurat@meta.data[[glue("{dbname}_2")]]

DefaultAssay(seurat) <- "RNA"

seurat$CellType <- droplevels(seurat$CellType)

convert <- c("ATL" = "LOH",  "DCT" = "DCT",
             "Endo" = "Endo",
             "ICA" = "Glom", "ICB" = "Glom", "Per" = "Glom",
             "Podo" = "Glom", "PC" = "Glom", "MC" = "Glom",
             "PTS1" = "PT", "PTS2" = "PT", "PTS3" = "PT", "PTS3T2" = "PT",
             "B lymph" = "Immune", "T lymph" = "Immune", "Macro" = "Immune",
             "Neutro" = "Immune", "NK" = "Immune"
             )

levels(seurat$CellType) <- 
  ifelse(levels(seurat$CellType) %in% names(convert),
         convert[levels(seurat$CellType)],
         levels(seurat$CellType))

Idents(seurat) <- seurat$CellType

DimPlot(seurat, split.by = "orig.ident", label = TRUE) +
  theme(legend.position = 'none')

saveRDS(seurat, here(dir, "rds", glue("WT_MT_ZK_seurat.rds")))

###########################
# Compute CellType markers
###########################
all_markers <- FindAllMarkers(seurat, only.pos = T)

write.table(all_markers, file = here(dir, "table", "all_markers.txt"),
            quote = FALSE, row.names = FALSE, sep = "\t")
```

## MT vs WT

The aim is to compute DEG of conditions MT vs WT in each cell types, and the enriched pathway of PT cells.

```{r}
library(here)
library(Seurat)
library(glue)
library(dplyr)
library(ggplot2)
source("fct_pathway.R")

###########################
# Compute DEGs for each CellType
###########################

treatment_sample <- "MT" 
base_sample <- "WT"

DEG <- list()

for (celltype in unique(seurat$CellType)){
  DEG[[celltype]] <- FindMarkers(seurat,
                                 ident.1 = treatment_sample,
                                 ident.2 = base_sample,
                                 group.by = "orig.ident",
                                 subset.ident = celltype)

  write.table(DEG[[celltype]] %>% tibble::rownames_to_column("gene"),
      file = here(dir, "table", "deg",
                 glue("deg_", 
                      gsub("\\W", "_", celltype),
                      "_{treatment_sample}vs{base_sample}.txt")),
      quote = FALSE, row.names = FALSE, sep = "\t")
}

###########################
# Compute pathways of DEGs
###########################

# Compute enrichment
OPT <- list()
OPT$comparison <- list()

term <- terms$All # terms is defined in fct_pathway.R file
celltype <- "PT"

# Compute ordered geneList
geneList <- DEG[[celltype]] %>%
  mutate(metric = avg_log2FC) %>%
  arrange(desc(metric)) %>%
  # deals with Inf numbers
  dplyr::mutate(metric = ifelse(is.finite(metric), metric,
                                .Machine$double.xmax * sign(metric))) %>%
  dplyr::select(metric) %>%
  tibble::rownames_to_column() %>%
  tibble::deframe()

OPT$comparison[[glue("GSEA {celltype} {term$name}")]] <- list(
  name = glue("GSEA {celltype} {term$name}"),
  term = terms[[term$name]],
  method = "GSEA",
  GSEA = list(geneList = geneList),
  pairwise_termsim = list(method = "JC")
)

OPT$comparison <- lapply(OPT$comparison, function(x){
  x$gsea_res <- pipeline_enrichment(x)
  return(x)
})

# Save results as tsv
lapply(OPT$comparison, function(x){
  file <- here(dir, "table", "pathway", glue("gsea_", gsub("\\W", "_", x$name), "_{treatment_sample}vs{base_sample}.tsv"))
  write.table(x$gsea_res, file,
              sep='\t', quote = FALSE, na = '', dec = ',', row.names = FALSE)
})

```

## ZK vs MT

The aim is to compute DEG of conditions ZK vs MT in each cell types, and the enriched pathway of PT cells.

```{r}
library(here)
library(Seurat)
library(glue)
library(dplyr)
library(ggplot2)
source("fct_pathway.R")

###########################
# Compute DEGs for each CellType
###########################

treatment_sample <- "ZK" 
base_sample <- "MT"

DEG <- list()

for (celltype in unique(seurat$CellType)){
  DEG[[celltype]] <- FindMarkers(seurat,
                                 ident.1 = treatment_sample,
                                 ident.2 = base_sample,
                                 group.by = "orig.ident",
                                 subset.ident = celltype)

  write.table(DEG[[celltype]] %>% tibble::rownames_to_column("gene"),
      file = here(dir, "table", "deg",
                 glue("deg_", 
                      gsub("\\W", "_", celltype),
                      "_{treatment_sample}vs{base_sample}.txt")),
      quote = FALSE, row.names = FALSE, sep = "\t")
}

###########################
# Compute pathways of DEGs
###########################

# Compute enrichment
OPT <- list()
OPT$comparison <- list()

term <- terms$All # terms is defined in script/fct_pathway.R file

celltype <- "PT"

# Compute ordered geneList
geneList <- DEG[[celltype]] %>%
  mutate(metric = avg_log2FC) %>%
  arrange(desc(metric)) %>%
  # deals with Inf numbers
  dplyr::mutate(metric = ifelse(is.finite(metric), metric,
                                .Machine$double.xmax * sign(metric))) %>%
  dplyr::select(metric) %>%
  tibble::rownames_to_column() %>%
  tibble::deframe()

OPT$comparison[[glue("GSEA {celltype} {term$name}")]] <- list(
  name = glue("GSEA {celltype} {term$name}"),
  term = terms[[term$name]],
  method = "GSEA",
  GSEA = list(geneList = geneList),
  pairwise_termsim = list(method = "JC")
)


OPT$comparison <- lapply(OPT$comparison, function(x){
  x$gsea_res <- pipeline_enrichment(x)
  return(x)
})

# Save results as tsv
lapply(OPT$comparison, function(x){
  file <- here(dir, "table", "pathway", glue("gsea_", gsub("\\W", "_", x$name), "_{treatment_sample}vs{base_sample}.tsv"))
  write.table(x$gsea_res, file,
              sep='\t', quote = FALSE, na = '', dec = ',', row.names = FALSE)
})

```

## CellChat

The aim is to run CellChat comparison analysis on MT and WT samples.

```{r}
library(CellChat)
library(future)
options(future.seed=TRUE)

seurat <- readRDS(here(dir, "rds", glue("WT_MT_ZK_seurat.rds")))

# Initialisation of db
CellChatDB <- CellChatDB.mouse # use CellChatDB.mouse if running on mouse data
showDatabaseCategory(CellChatDB)
CellChatDB.use <- subsetDB(CellChatDB, search = "Secreted Signaling", key = "annotation") 

# Create individual WT and MT CellChat objects
samples <- list("WT" = "WT", "MT" = "MT")

cellchat_list <- lapply(samples, function(x){
  # Initialisation of seurat object
  seurat.subset <- subset(seurat, subset = orig.ident == x)

  data.input <- seurat.subset[["RNA"]]$data # normalized data matrix
  # For Seurat version >= “5.0.0”, get the normalized data via `seurat_object[["RNA"]]$data`
  labels <- Idents(seurat.subset)
  meta <- data.frame(group = labels, row.names = names(labels)) # create a dataframe of the cell labels

  cellchat <- createCellChat(object = seurat.subset, group.by = "CellType", assay = "RNA")

  #cellchat <- addMeta(cellchat, meta = meta, meta.name = "labels")
  #cellchat <- setIdent(cellchat, ident.use = "labels") # set "labels" as default cell identity
  #levels(cellchat@idents) # show factor levels of the cell labels
  groupSize <- as.numeric(table(cellchat@idents))

  cellchat@DB <- CellChatDB.use

  # subset the expression data of signaling genes for saving computation cost
  cellchat <- subsetData(cellchat) # This step is necessary even if using the whole database

  future::plan("multisession", workers = 4) # do parallel
  options(future.seed=TRUE)

  cellchat <- identifyOverExpressedGenes(cellchat, do.fast = FALSE,
                                         thresh.pc = 0.1, thresh.fc = 0.1)
  cellchat <- identifyOverExpressedInteractions(cellchat)
  cellchat <- computeCommunProb(cellchat, type = "triMean")

  cellchat <- filterCommunication(cellchat, min.cells = 10)

  # Infer the cell-cell communication at a signaling pathway level
  cellchat <- computeCommunProbPathway(cellchat)

  # Calculate the aggregated cell-cell communication network
  cellchat <- aggregateNet(cellchat)
  cellchat <- netAnalysis_computeCentrality(cellchat)
  return(cellchat)
})

# Extract the inferred cellular communication network as a data frame
df.net <- subsetCommunication(cellchat_list$WT)
DT::datatable(df.net)

# Extract the inferred cellular communication network as a data frame
df.net <- subsetCommunication(cellchat_list$MT)
DT::datatable(df.net)

# Merge objects
cellchat <- mergeCellChat(cellchat_list, add.names = names(cellchat_list))

cellchat@netP$WT$pathways
cellchat@netP$MT$pathways

#  Circle plot showing differential number of interactions or interaction strength among different cell populations across two datasets
netVisual_diffInteraction(cellchat, weight.scale = T, measure = "count")

# Circle plot showing the number of interactions or interaction strength among different cell populations across multiple datasets
weight.max <- getMaxWeight(cellchat_list, attribute = c("idents","count"))
for (i in 1:length(cellchat_list)) {
  netVisual_circle(cellchat_list[[i]]@net$count, weight.scale = T, label.edge= T, edge.weight.max = weight.max[2], edge.width.max = 12)
  title(paste0("Number of interactions - ", names(cellchat_list)[i]))
}

saveRDS(cellchat, file = here(dir, "rds", "cellchat_merged.rds"))
```
