---
title: "scRNA_seq"
format: html
toc: true
---

# Set up

```{r}
library(here)
library(glue)
library(Seurat)
library(ggplot2)
library(dplyr)
library(HGNChelper)
library(openxlsx)
library(Seurat)
library(DoubletFinder)
source(here("utils/sctype_wrapper.R"))
source(here("utils/fct_pathway.R"))
# For CellChat
library(CellChat)
library(future)
options(future.seed=TRUE)

dir <- "scRNA_seq/"
result <- "result/"

# Create directories
directories <- c(here(dir, result, "rds"), here(dir, result, "figure/cellchat"),here(dir, result, "figure/seurat"), here(dir, result, "table"))
sapply(directories, function(x){dir.create(x, showWarnings = FALSE, recursive = TRUE)})

```

## Load data from cellbender

CellBender remove cells with ambiant RNA (cf https://www.10xgenomics.com/analysis-guides/introduction-to-ambient-rna-correction).

```{r}
list_seurat <- list(
  # WT
  "WT" = list(
    condition = "WT",
    h5.file = here("result/260206_cellbender/MNGC484_cellbender_seurat.h5")),
  # Cyp24a1 null 10 weeks
  "MT" = list(
    condition = "MT",
    h5.file = here("result/260206_cellbender/MNGC485_cellbender_seurat.h5")),
  # ZK-treated Cyp24a1 null 10 weeks
  "ZK" = list(
    condition = "ZK",
    h5.file = here("result/260206_cellbender/MNGC486_cellbender_seurat.h5"))
)

# Creation of Seurat objects
list_seurat <- lapply(list_seurat, function(x){
  # Load the data
  data <- Read10X_h5(x$h5.file)
  seurat <- CreateSeuratObject(data,
                               project = x$condition,
                               assay = "RNA",
                               min.cells = 10,
                               min.features = 100,
                               names.field = 1,
                               names.delim = "_",
                               meta.data = NULL
  )
  sample <- unique(seurat$orig.ident)

  # Add percent.mt feature
  feature <- list(col.name = "percent.mt",
                  pattern = "^MT-|^mt-|^Mt-")

  seurat <- PercentageFeatureSet(object = seurat,
                                 col.name = feature$col.name,
                                 pattern = feature$pattern)

  feature <- list(col.name = "percent.ribo",
                  pattern = "^Rp[sl][[:digit:]]|^Rp[[:digit:]]|^Rpsa")

  seurat <- PercentageFeatureSet(object = seurat,
                                 col.name = feature$col.name,
                                 pattern = feature$pattern)

  # Save VlnPlot of unfiltered data
  VlnPlot(seurat, features = c("nFeature_RNA", "nCount_RNA", 
                               "percent.mt", "percent.ribo"),
          pt.size = 0.1, ncol = 2)
  ggsave(here(dir, result, glue("figure/seurat/{sample}_vln_before_filter.png")))

  # Filter data
  seurat <- subset(seurat, subset = nFeature_RNA > 100 &
                     nFeature_RNA < 7000 &
                     percent.mt < 50)

  # Save VlnPlot of filtered data
  VlnPlot(seurat, features = c("nFeature_RNA", "nCount_RNA", 
                               "percent.mt", "percent.ribo"),
          pt.size = 0.1, ncol = 2)
  ggsave(here(dir, result, glue("figure/seurat/{sample}_vln_after_filter.png")))
  
  return(seurat)
})

saveRDS(list_seurat, here(dir, result, "rds", "list_seurat_filter.rds"))
```

## DoubletFinder

DoubletFinder should be run on individual objects, not integrated ones: https://github.com/chris-mcginnis-ucsf/DoubletFinder/issues/101

```{r}
for (x in list_seurat){
  sample <- unique(x$orig.ident)
  x <- NormalizeData(x)
  x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 2000)
  x <- ScaleData(x)
  x <- RunPCA(x)
  ElbowPlot(x, ndims=50)
  ggsave(here(dir, result, glue("figure/seurat/{sample}_elbowplot.png")))

  x <- RunUMAP(x, dims = 1:30)
  x <- FindNeighbors(x, dims = 1:30)
  x <- FindClusters(x, resolution =  0.8)
  
  DimPlot(x, reduction="umap", 
          group.by=c("RNA_snn_res.0.8"))
  ggsave(here(dir, result, glue("figure/seurat/{sample}_umap_before_doubletfinder.png")))
  
  ## pK Identification (no ground-truth) ---------------------------------------------------------------------------------------
  sweep.res<- paramSweep(x, PCs = 1:30, sct = FALSE)
  sweep.stat <- summarizeSweep(sweep.res, GT = FALSE)
  bcmvn <- find.pK(sweep.stat)
  
  ggplot(bcmvn, aes(x = pK, y = BCmetric)) +
      geom_point() +
      theme_bw() + 
      ggtitle("BC metric vs pK (choose pK with max)") + 
      theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  ggsave(here(dir, result, glue("figure/seurat/{sample}_doubletfinder_bcmvn.png")))

  pK = as.numeric(as.character((bcmvn[which(bcmvn$BCmetric == max(bcmvn$BCmetric)),]$pK)))
  
  ## Homotypic Doublet Proportion Estimate -------------------------------------------------------------------------------------
  homotypic.prop <- modelHomotypic(x$RNA_snn_res.0.8)           
  nExp_poi <- round(0.05*nrow(x@meta.data))  
  nExp_poi.adj <- round(nExp_poi*(1-homotypic.prop))
  
  ## Run DoubletFinder with varying classification stringencies ----------------------------------------------------------------
  x <- doubletFinder(x, PCs = 1:30, pN = 0.25, pK = pK, nExp = nExp_poi, reuse.pANN = NULL, sct = FALSE)
  reuse.pANN <- colnames(x@meta.data)[length(colnames(x@meta.data))-1]
  DF.classifications.nExp <- colnames(x@meta.data)[length(colnames(x@meta.data))]
  x <- doubletFinder(x, PCs = 1:30, pN = 0.25, pK = pK, nExp = nExp_poi.adj, reuse.pANN = reuse.pANN, sct = FALSE)
  DF.classifications.nExp.adj <- colnames(x@meta.data)[length(colnames(x@meta.data))]
  
  p1 <- DimPlot(x, split.by = "orig.ident", 
          group.by = c(DF.classifications.nExp, DF.classifications.nExp.adj), reduction = "umap") 
  p2 <- FeaturePlot(x, split.by = "orig.ident" , reduction = "umap", 
              feature=c("percent.mt", "percent.ribo", "nCount_RNA"))
  
  patchwork::wrap_plots(p1, p2, ncol=2)
  ggsave(here(dir, result, glue("figure/seurat/{sample}_umap_doubletfinder.png")))

  x <- subset(x, !!sym(DF.classifications.nExp.adj) == "Singlet")
  
  DimPlot(x, reduction="umap", 
          group.by=c("RNA_snn_res.0.8"))
  ggsave(here(dir, result, glue("figure/seurat/{sample}_umap_after_doubletfinder.png")))
  
  list_seurat[[as.character(sample)]] <- x
}

saveRDS(list_seurat, here(dir, result, "rds", "list_seurat_doubletfinder.rds"))
```

## Integration of WT, MT and ZK objects

```{r}
# Integration of WT, MT and ZK objects

# Creation of the combined Seurat object
seurat <- merge(x = list_seurat$WT,
                y = list_seurat[c("MT", "ZK")])
# Normalize each object
seurat <- NormalizeData(seurat)
seurat <- FindVariableFeatures(seurat,
                               selection.method = "vst",
                               nfeatures = 2000)

top10 <- head(VariableFeatures(seurat), 10)
plot1 <- VariableFeaturePlot(seurat)
LabelPoints(plot = plot1, points = top10, repel = FALSE)

seurat <- ScaleData(seurat, verbose = F)
seurat <- RunPCA(seurat, verbose = F)
ElbowPlot(seurat, ndims=50)
ggsave(here(dir, result, glue("figure/seurat/integrated_elbowplot.png")))

# Integrate the datasets
seurat <- IntegrateLayers(seurat, method = CCAIntegration,
                          orig = "pca", new.reduction = "integrated.cca",
                          verbose = T)
# Cluster the integrated data
seurat <- FindNeighbors(seurat, reduction = "integrated.cca", dims = 1:30)
seurat <- FindClusters(seurat, resolution =  0.8)

seurat <- RunUMAP(seurat, reduction = "integrated.cca",
                  dims = 1:30, return.model = T, verbose = F)

# Create common counts and data layers
# seurat <- JoinLayers(seurat)

DimPlot(seurat, group.by = "orig.ident", reduction = "umap")
ggsave(here(dir, result, glue("figure/seurat/integrated_umap.png")))

DimPlot(seurat, split.by = "orig.ident", 
        group.by = "seurat_clusters", reduction = "umap", label=T)
ggsave(here(dir, result, glue("figure/seurat/integrated_umap_split.png")), width=10)


# Save VlnPlot of unfiltered data
VlnPlot(seurat, features = c("nFeature_RNA", "nCount_RNA", 
                             "percent.mt", "percent.ribo"),
        group.by = "seurat_clusters",
        pt.size = 0.1, ncol = 2)
ggsave(here(dir, result, glue("figure/seurat/integrated_vln_cluster_qual.png")), width=10)

DefaultAssay(seurat) <- "RNA"
saveRDS(seurat, here(dir, result, "rds", "WT_MT_ZK_seurat_integrated.rds"))
```


## Reannotate
 
### Global annotation with sctype

```{r}
tissue = "Kidney"

# Create marker table in sctype format 
create_markers_table <- FALSE # Do it only once
if (create_markers_table){
  markers_all <- read.xlsx(here(dir,"data/1-s2.0-S2589004223009549-mmc5.xlsx"))
  markers <- markers_all %>%
    group_by(Cell.Population) %>%
    slice_head(n = 10) %>%
    aggregate(Gene.symbol ~ Cell.Population,
              FUN = function(x) paste(x, collapse = ",")) %>%
    mutate(tissueType = "Kidney", .before = 1) %>%
    rename("cellName" = "Cell.Population",
           "geneSymbolmore1" = "Gene.symbol") %>%
    mutate(geneSymbolmore2 = NA,
           shortName = cellName,
           .after = "geneSymbolmore1")
  write.xlsx(markers, file = "data/kidney_db.xlsx",
             sheetName="Kidney", append=TRUE)
}

# Compute sctype annotation
dbname <- "kidney_db"
kidney_db <- here(dir, glue("data/{dbname}.xlsx"))

seurat <- run_sctype(seurat, assay = "RNA", scaled = TRUE,
                     known_tissue_type = tissue,
                     custom_marker_file = kidney_db,
                     name = dbname)
seurat@meta.data[[dbname]] <- seurat@meta.data[[dbname]] %>% as.factor()

DimPlot(seurat, group.by = dbname,
        split.by = "orig.ident", label = TRUE) +
  theme(legend.position = 'none')
ggsave(here(dir, result, glue("figure/seurat/annotation_v1.png")), width=10)

# Number of cells per annotation
summary(seurat@meta.data[[dbname]] %>% as.factor())

# Remove Unknown and re-annotate
expr <- FetchData(seurat, vars = dbname)
seurat <- seurat[, which(expr != "Unknown")]

# Re-run seurat analysis
seurat <- RunPCA(seurat, npcs = 50, verbose = T)
ElbowPlot(seurat, ndims = 50)
ggsave(here(dir, result, glue("figure/seurat/annotation_elbowplot.png")))
 
seurat <- IntegrateLayers(object = seurat, method = CCAIntegration,
                        orig.reduction = "pca", new.reduction = "integrated.cca",
                        verbose = FALSE)

# Re-join layers after integration
seurat[["RNA"]] <- JoinLayers(seurat[["RNA"]])
# to Re-split run:
# seurat_kidneymarkers[["RNA"]] <- split(seurat_kidneymarkers[["RNA"]], f = seurat_kidneymarkers$orig.ident)

seurat <- FindNeighbors(seurat,reduction = "integrated.cca", dims = 1:30)
seurat <- FindClusters(seurat, resolution =  0.8)
seurat <- RunUMAP(seurat, dims = 1:30, reduction = "integrated.cca")

# Re-compute sctype annotation
seurat <- run_sctype(seurat, assay = "RNA",
                             scaled = TRUE,
                              known_tissue_type = tissue,
                              custom_marker_file = kidney_db,
                              name = glue("{dbname}_2"))
seurat@meta.data[[glue("{dbname}_2")]] <- seurat@meta.data[[glue("{dbname}_2")]] %>% as.factor()

# Number of cells per annotation
summary(seurat@meta.data[[glue("{dbname}_2")]] %>% as.factor())

DimPlot(seurat, group.by = glue("{dbname}_2"),
        split.by = "orig.ident", label = TRUE) +
  theme(legend.position = 'none')
ggsave(here(dir, result, glue("figure/seurat/annotation_v2.png")), width=10)

# Remove Unknown 
expr <- FetchData(seurat, vars = glue("{dbname}_2"))
seurat <- seurat[, which(expr != "Unknown")]

DimPlot(seurat, group.by = glue("{dbname}_2"),
        split.by = "orig.ident", label = TRUE) +
  theme(legend.position = 'none')
ggsave(here(dir, result, glue("figure/seurat/annotation_v2_no_unknown.png")), width=10)

DimPlot(seurat, group.by = glue("seurat_clusters"),
        split.by = "orig.ident", label = TRUE) +
  theme(legend.position = 'none')
ggsave(here(dir, result, glue("figure/seurat/annotation_v2_seurat_clusters.png")), width=10)

# Save results
saveRDS(seurat,
        here(dir, result, "rds", "WT_MT_ZK_seurat_annotated.rds"))
```

### CellType reannotation based on clusters

```{r}
seurat$CellType <- seurat@meta.data[[glue("seurat_clusters")]]

seurat$CellType <- droplevels(seurat$CellType)

convert <- c("12" = "T lymph", 
             "20" = "B lymph", 
             "27" = "Macro", 
             "8" = "Macro", 
             "30" = "Macro", 
             "19" = "Macro", 
             "29" = "Macro", 
             "28" = "Neutro", 
             "26" = "Endo", 
             "0" = "Endo", 
             "25" = "ICA/ICB", 
             "18" = "CNT/DCT", 
             "24" = "LOH", 
             "16" = "LOH", 
             "15" = "LOH", 
             "23" = "LOH",
             "14" = "CDPC", 
             "1" = "PTS1", 
             "2" = "PTS3", 
             "3" = "PTS3", 
             "4" = "PTS3", 
             "7" = "PTS3", 
             "9" = "PTS3", 
             "10" = "PTS2", 
             "11" = "PTS2", 
             "5" = "Profibrotic", 
             "6" = "Profibrotic", 
             "17" = "Profibrotic", 
             "22" = "Profibrotic", 
             "21" = "PTS3")

levels(seurat$CellType) <- 
  ifelse(levels(seurat$CellType) %in% names(convert),
         convert[levels(seurat$CellType)],
         levels(seurat$CellType))

seurat$CellType<- factor(seurat$CellType, 
                         levels = c("PTS1", "PTS2", "PTS3", "Profibrotic", 
                                    "LOH", "ICA/ICB", "CNT/DCT", "CDPC", "Endo",
                                    "B lymph", "T lymph", "Macro", "Neutro"))

seurat$orig.ident <- factor(seurat$orig.ident, levels = c("WT", "MT", "ZK"))

DimPlot(seurat, group.by = glue("CellType"),
        split.by = "orig.ident", label = TRUE) +
  theme(legend.position = 'none')
ggsave(here(dir, result, glue("figure/seurat/dimplot_CellType_label.png")), width=10)

DimPlot(seurat, group.by = glue("CellType"),
        split.by = "orig.ident", label = FALSE) 
ggsave(here(dir, result, glue("figure/seurat/dimplot_CellType_nolabel.png")), width=10)
```

### CellType_2 with PT grouped

```{r}
seurat$CellType_2 <- seurat@meta.data[[glue("CellType")]]

seurat$CellType_2 <- droplevels(seurat$CellType_2)

convert <- c("PTS1" = "PT", 
             "PTS2" = "PT", 
             "PTS3" = "PT", 
             "Profibrotic" = "PT")

levels(seurat$CellType_2) <- 
  ifelse(levels(seurat$CellType_2) %in% names(convert),
         convert[levels(seurat$CellType_2)],
         levels(seurat$CellType_2))

seurat$CellType_2<- factor(seurat$CellType_2, 
                         levels = c("PT", "LOH", "ICA/ICB", "CNT/DCT", "CDPC", "Endo", "B lymph", "T lymph", "Macro", "Neutro"))

DimPlot(seurat, group.by = glue("CellType_2"),
        split.by = "orig.ident", label = TRUE) +
  theme(legend.position = 'none')
ggsave(here(dir, result, glue("figure/seurat/dimplot_CellType_2_label.png")), width=10)

DimPlot(seurat, group.by = glue("CellType_2"),
        split.by = "orig.ident", label = FALSE) 
ggsave(here(dir, result, glue("figure/seurat/dimplot_CellType_2_nolabel.png")), width=10)
```

### Save final seurat object

```{r}
# Save results
saveRDS(seurat,
        here(dir, result, "rds", "WT_MT_ZK_seurat.rds"))
```

## Cell type markers

The aim is to compute cell type markers.

```{r}
# seurat <- readRDS(here(dir, result, "rds", glue("WT_MT_ZK_seurat.rds")))
dbname <- "kidney_db"

###########################
# Compute seurat_clusters markers
###########################
Idents(seurat) <- "seurat_clusters"
all_markers <- FindAllMarkers(seurat, only.pos = T)

write.table(all_markers, file = here(dir, result, "table", glue("all_markers_seurat_clusters.txt")),
            quote = FALSE, row.names = FALSE, sep = "\t")

###########################
# Compute dbname_2 markers
###########################
Idents(seurat) <- glue("{dbname}_2")
all_markers <- FindAllMarkers(seurat, only.pos = T)

write.table(all_markers, file = here(dir, result, "table", glue("all_markers_{dbname}_2.txt")),
            quote = FALSE, row.names = FALSE, sep = "\t")

###########################
# Compute CellType markers
###########################
Idents(seurat) <- "CellType"
all_markers <- FindAllMarkers(seurat, only.pos = T)

write.table(all_markers, file = here(dir, result, "table", "all_markers_CellType.txt"),
            quote = FALSE, row.names = FALSE, sep = "\t")

###########################
# Compute CellType_2 markers
###########################
Idents(seurat) <- "CellType_2"
all_markers <- FindAllMarkers(seurat, only.pos = T)

write.table(all_markers, file = here(dir, result, "table", "all_markers_CellType_2.txt"),
            quote = FALSE, row.names = FALSE, sep = "\t")

```

## Dotplot of celltype markers


```{r}
markers <- c("Slc34a1", "Slc25a21",
             "Ttc36", "Nhs", "Umod", 
             "Atp6v1g3", "Clu", "Slc12a3",
             "Emcn", "Cd79a" , "Cd3d", "Cd74", "S100a8")

DotPlot(seurat, group.by= glue("CellType"), 
        features=markers, dot.min = 0.1, col.min=0) + RotatedAxis() 
ggsave(here(dir, result, glue("figure/seurat/dotplot_CellType_markers.png")), width=8, height=5)
```

## DEG MT vs WT 

The aim is to compute DEG of conditions MT vs WT in each cell types, and the enriched pathway of PT cells.

```{r}
###########################
# Compute DEGs for each CellType
###########################

metadata_key <- "CellType"
Idents(seurat) <- seurat@meta.data[[metadata_key]]
treatment_sample <- "MT" 
base_sample <- "WT"

DEG <- list()

deg_dir <- here(dir, result, "table", glue("deg_{metadata_key}_MTvsWT"))
dir.create(deg_dir, showWarnings = FALSE, recursive = TRUE)

for (celltype in unique(seurat@meta.data[[metadata_key]])){
  DEG[[celltype]] <- FindMarkers(seurat,
                                 ident.1 = treatment_sample,
                                 ident.2 = base_sample,
                                 group.by = "orig.ident",
                                 subset.ident = celltype)

  write.table(DEG[[celltype]] %>% tibble::rownames_to_column("gene"),
      file = here(deg_dir,
                 glue("deg_",
                      gsub("\\W", "_", celltype),
                      "_{treatment_sample}vs{base_sample}.txt")),
      quote = FALSE, row.names = FALSE, sep = "\t")
}

###########################
# Compute pathways of DEGs
###########################

# Compute enrichment
OPT <- list()
OPT$comparison <- list()

term <- terms$All # terms is defined in fct_pathway.R file

for (celltype in unique(seurat@meta.data[[metadata_key]])){
  # Compute ordered geneList
  geneList <- DEG[[celltype]] %>%
    mutate(metric = avg_log2FC) %>%
    arrange(desc(metric)) %>%
    # deals with Inf numbers
    dplyr::mutate(metric = ifelse(is.finite(metric), metric,
                                  .Machine$double.xmax * sign(metric))) %>%
    dplyr::select(metric) %>%
    tibble::rownames_to_column() %>%
    tibble::deframe()
  
  OPT$comparison[[glue("GSEA {celltype} {term$name}")]] <- list(
    name = glue("GSEA {celltype} {term$name}"),
    term = terms[[term$name]],
    method = "GSEA",
    GSEA = list(geneList = geneList, pvalueCutoff = 1),
    pairwise_termsim = list(method = "JC")
  )
}

OPT$comparison <- lapply(OPT$comparison, function(x){
  x$gsea_res <- pipeline_enrichment(x)
  return(x)
})

# Save results as tsv
pathway_dir <- here(dir, result, "table", glue("pathway_{metadata_key}_MTvsWT"))
dir.create(pathway_dir, showWarnings = FALSE, recursive = TRUE)

lapply(OPT$comparison, function(x){
  file <- here(pathway_dir, glue("gsea_", gsub("\\W", "_", x$name), "_{treatment_sample}vs{base_sample}_pvalcutoff_1.tsv"))
  write.table(x$gsea_res, file,
              sep='\t', quote = FALSE, na = '', dec = ',', row.names = FALSE)
})

```

## DEG ZK vs MT

The aim is to compute DEG of conditions ZK vs MT in each cell types, and the enriched pathway of PT cells.

```{r}
###########################
# Compute DEGs for each CellType
###########################

metadata_key <- "CellType_2"
Idents(seurat) <- seurat@meta.data[[metadata_key]]
treatment_sample <- "ZK" 
base_sample <- "MT"

DEG <- list()
deg_dir <- here(dir, result, "table", glue("deg_{metadata_key}_ZKvsMT"))
dir.create(deg_dir, showWarnings = FALSE, recursive = TRUE)


for (celltype in unique(seurat@meta.data[[metadata_key]])){
  DEG[[celltype]] <- FindMarkers(seurat,
                                 ident.1 = treatment_sample,
                                 ident.2 = base_sample,
                                 group.by = "orig.ident",
                                 subset.ident = celltype)

  write.table(DEG[[celltype]] %>% tibble::rownames_to_column("gene"),
      file = here(deg_dir,
                 glue("deg_",
                      gsub("\\W", "_", celltype),
                      "_{treatment_sample}vs{base_sample}.txt")),
      quote = FALSE, row.names = FALSE, sep = "\t")
}

###########################
# Compute pathways of DEGs
###########################

# Compute enrichment
OPT <- list()
OPT$comparison <- list()

term <- terms$All # terms is defined in script/fct_pathway.R file


for (celltype in unique(seurat@meta.data[[metadata_key]])){
  # Compute ordered geneList
  geneList <- DEG[[celltype]] %>%
    mutate(metric = avg_log2FC) %>%
    arrange(desc(metric)) %>%
    # deals with Inf numbers
    dplyr::mutate(metric = ifelse(is.finite(metric), metric,
                                  .Machine$double.xmax * sign(metric))) %>%
    dplyr::select(metric) %>%
    tibble::rownames_to_column() %>%
    tibble::deframe()
  
  OPT$comparison[[glue("GSEA {celltype} {term$name}")]] <- list(
    name = glue("GSEA {celltype} {term$name}"),
    term = terms[[term$name]],
    method = "GSEA",
    GSEA = list(geneList = geneList, pvalueCutoff = 1),
    pairwise_termsim = list(method = "JC")
  )

}


OPT$comparison <- lapply(OPT$comparison, function(x){
  x$gsea_res <- pipeline_enrichment(x)
  return(x)
})

# Save results as tsv
pathway_dir <- here(dir, result, "table", glue("pathway_{metadata_key}_ZKvsMT"))
dir.create(pathway_dir, showWarnings = FALSE, recursive = TRUE)

lapply(OPT$comparison, function(x){
  file <- here(pathway_dir, glue("gsea_", gsub("\\W", "_", x$name), "_{treatment_sample}vs{base_sample}_pvalcutoff_1.tsv"))
  write.table(x$gsea_res, file,
              sep='\t', quote = FALSE, na = '', dec = ',', row.names = FALSE)
})

```

## CellChat

The aim is to run CellChat comparison analysis on MT and WT samples.

```{r}
seurat <- readRDS(here(dir, result, "rds", glue("WT_MT_ZK_seurat.rds")))
metadata_key <- "CellType_2"

# Initialisation of db
CellChatDB <- CellChatDB.mouse # use CellChatDB.mouse if running on mouse data
showDatabaseCategory(CellChatDB)
CellChatDB.use <- subsetDB(CellChatDB, search = "Secreted Signaling", key = "annotation") 

# Create individual WT and MT CellChat objects
samples <- list("WT" = "WT", "MT" = "MT")

cellchat_list <- lapply(samples, function(x){
  # Initialisation of seurat object
  seurat.subset <- subset(seurat, subset = orig.ident == x)

  data.input <- seurat.subset[["RNA"]]$data # normalized data matrix
  # For Seurat version >= “5.0.0”, get the normalized data via `seurat_object[["RNA"]]$data`
  labels <- Idents(seurat.subset)
  meta <- data.frame(group = labels, row.names = names(labels)) # create a dataframe of the cell labels

  cellchat <- createCellChat(object = seurat.subset, group.by = metadata_key, assay = "RNA")

  #cellchat <- addMeta(cellchat, meta = meta, meta.name = "labels")
  #cellchat <- setIdent(cellchat, ident.use = "labels") # set "labels" as default cell identity
  #levels(cellchat@idents) # show factor levels of the cell labels
  groupSize <- as.numeric(table(cellchat@idents))

  cellchat@DB <- CellChatDB.use

  # subset the expression data of signaling genes for saving computation cost
  cellchat <- subsetData(cellchat) # This step is necessary even if using the whole database

  future::plan("multisession", workers = 4) # do parallel
  options(future.seed=TRUE)

  cellchat <- identifyOverExpressedGenes(cellchat, do.fast = FALSE,
                                         thresh.pc = 0.1, thresh.fc = 0.1)
  cellchat <- identifyOverExpressedInteractions(cellchat)
  cellchat <- computeCommunProb(cellchat, type = "triMean")

  cellchat <- filterCommunication(cellchat, min.cells = 10)

  # Infer the cell-cell communication at a signaling pathway level
  cellchat <- computeCommunProbPathway(cellchat)

  # Calculate the aggregated cell-cell communication network
  cellchat <- aggregateNet(cellchat)
  cellchat <- netAnalysis_computeCentrality(cellchat)
  return(cellchat)
})

# Extract the inferred cellular communication network as a data frame
df.net <- subsetCommunication(cellchat_list$WT)
DT::datatable(df.net)

# Extract the inferred cellular communication network as a data frame
df.net <- subsetCommunication(cellchat_list$MT)
DT::datatable(df.net)

# Merge objects
cellchat <- mergeCellChat(cellchat_list, add.names = names(cellchat_list))

cellchat@netP$WT$pathways
cellchat@netP$MT$pathways

#  Circle plot showing differential number of interactions or interaction strength among different cell populations across two datasets
netVisual_diffInteraction(cellchat, weight.scale = T, measure = "count")

# Circle plot showing the number of interactions or interaction strength among different cell populations across multiple datasets
weight.max <- getMaxWeight(cellchat_list, attribute = c("idents","count"))
for (i in 1:length(cellchat_list)) {
  pdf(here(dir, result, glue("figure/cellchat/nb_interaction_{names(cellchat_list)[i]}_{metadata_key}.pdf")))
  plot.new()
  netVisual_circle(cellchat_list[[i]]@net$count, weight.scale = T, label.edge= T, edge.weight.max = weight.max[2], edge.width.max = 12)
  title(paste0("Number of interactions - ", names(cellchat_list)[i]))
  dev.off()
}

saveRDS(cellchat, file = here(dir, result, "rds", "cellchat_merged.rds"))
```


# `sessionInfo()`

```{r}
sessionInfo()
```


